\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax
\chapter*{Preface}

In late 2013 I came back to Aalto University to continue my studies after a hiatus of five years. While going through the courses, I stumbled upon the new computer graphics courses by Prof. Jaakko Lehtinen. They immediately piqued my interest as the content and course execution seemed very well done. I started to think that this could be my ``exit strategy'' for my very long overdue studies. Realizing that the course assignments did not have any upper bounds on points awarded (and some nice prizes), I brushed up my graphics maths and went to town with the exercises. After getting quite high scores, Jaakko promised that I could join the graphics group and do my thesis there.

In summer 2016 we started brainstorming different ideas for my final project. I wanted to do something practical while including GPU computing, rendering, and neural networks. Ultimately, we ended up with the subject of this thesis; trying to track human facial geometry densely using fully convolutional neural networks. At that time there were no publications of the exact method we were going to use. Not having too much background in machine learning, I spent the fall of 2016 studying it and familiarizing myself with the various software needed. The better part of 2017 we spent iteratively improving the network model, training data, and augmentations until we had the results we were looking for. In the end, this work contains some math and theory, but even more practical software engineering and exciting visuals. Just how I like it!

I would like to thank Prof. Jaakko Lehtinen and D.Sc. Miika Aittala for advising the thesis and keeping my to-do list full, Prof. Ville Kyrki for supervising, Markus Kettunen and Ari Silvennoinen for fruitful discussions at the office, Timo Aila from NVIDIA Research for good ideas and advice, Remedy Entertainment for models and textures, team Luolamiehet for abundant help during my general studies, remedial massage therapist Timi Hilakari for keeping me physically able to write, and last but not least my family (Jyrki, Maarit, Jenni, Toini, Risto) for supporting and keeping a roof over my head.

\bigskip

The calculations presented in this thesis were performed using computer resources within the Aalto University School of Science "Science-IT" project.

\bigskip

The author wishes to acknowledge CSC – IT Center for Science, Finland, for generous computational resources.

\bigskip

\noindent Espoo, 20.11.2017

\noindent Mikko Ronkainen

\endgroup
